Main Analyses of the Combined Dataset

### 2021-05-10 manuscript/manuscript.tex
species range corrections, fig 2 tweaks

\documentclass[11pt]{article}
\RequirePackage{fullpage}
\RequirePackage[font=small,labelfont=bf]{caption}
\RequirePackage{amsmath,amssymb,amsthm}
\RequirePackage{mathtools}
\RequirePackage{graphicx}
\RequirePackage[normalem]{ulem}
\RequirePackage[hidelinks]{hyperref}
\RequirePackage{wrapfig}
\RequirePackage{subcaption}
\RequirePackage{authblk}
\RequirePackage{bm}
\RequirePackage{bbm}
\RequirePackage{tikz}
\RequirePackage{wrapfig}

@@

>>>>

\documentclass[11pt]{article}
\RequirePackage{fullpage}
\RequirePackage[font=small,labelfont=bf]{caption}
\RequirePackage{amsmath,amssymb,amsthm}
\RequirePackage{mathtools}
\RequirePackage{tabularx}
\RequirePackage{graphicx}
\RequirePackage[normalem]{ulem}
\RequirePackage[hidelinks]{hyperref}
\RequirePackage{wrapfig}
\RequirePackage{subcaption}
\RequirePackage{authblk}
\RequirePackage{bm}
\RequirePackage{bbm}
\RequirePackage{tikz}
\RequirePackage{wrapfig}

@@

The approximate census sizes I have estimated here

\begin{figure}[!htb]
  \centering
  \includegraphics[]{figures/species_counts.pdf}

  \label{suppfig:species_counts}
\end{figure}

\begin{table}[ht]
  \resizebox{\textwidth}{!}{%
\centering
\begin{tabular}{lr|rr|rrrr|r}
  &  & \multicolumn{2}{c}{\bf Bar-On et al.}  & \multicolumn{4}{c}{\bf Present study}  &  \\
  phylum & total species ($T$) & biomass ($B$) & prop. biomass & biomass ($b$) &  prop. biomass & num. species ($n$) &prop. total species ($f = \nicefrac{n}{T}$) & factor ($\nicefrac{b}{f B}$) \\
  \hline
  Arthropoda & $1.26 \times 10^{6}$ & 1.20 & 0.4635 & $2.77 \times 10^{-4}$ & 0.0100 &  68 & $5.41 \times 10^{-5}$ & 4.26 \\
  Chordata & $5.41 \times 10^{4}$ & 0.87 & 0.3357 & $2.68 \times 10^{-2}$ & 0.9720 &  68 & $1.26 \times 10^{-3}$ & 24.50 \\
  Annelida & $1.70 \times 10^{4}$ & 0.20 & 0.0772 & $1.24 \times 10^{-5}$ & 0.0005 &   3 & $1.76 \times 10^{-4}$ & 0.35 \\
  Mollusca & $9.54 \times 10^{4}$ & 0.20 & 0.0772 & $4.47 \times 10^{-4}$ & 0.0162 &  13 & $1.36 \times 10^{-4}$ & 16.40 \\
  Cnidaria & $1.60 \times 10^{4}$ & 0.10 & 0.0386 & $3.21 \times 10^{-5}$ & 0.0012 &   2 & $1.25 \times 10^{-4}$ & 2.57 \\
  Nematoda & $2.50 \times 10^{4}$ & 0.02 & 0.0077 & $4.00 \times 10^{-6}$ & 0.0001 &   1 & $4.00 \times 10^{-5}$ & 5.00 \\
\end{tabular}}

  \caption{How the total carbon biomass estimates by phylum from
    \textcite{Bar-On2018-kc} compare to the implied biomass estimates from this
    study. All biomass estimates are carbon biomass, and the proportions are of
    total biomass with respect to the study. The proportion of biomass in this
    study compared to the Bar-On et al. estimates \textcite{Bar-On2018-kc}
    indicates chordates are over-represented and arthropods are
    under-represented in the present study. Total species by phylum estimates
  are from
\textcite{Reaka-Kudla1996-fe,Nicol1969-qx,Zhang2013-zc,Chapman2009-fw}. The
ratio column is the ratio of total biomass implied by the $N_c$ estimates of
each species in a phylum to the actual biomass of that phylum.}

\end{table}

### 2021-05-10 notebooks/main_analysis.Rmd
species range corrections, fig 2 tweaks

# sources, 
# https://www.environment.gov.au/system/files/pages/2ee3f4a1-f130-465b-9c7a-79373680a067/files/nlsaw-2nd-complete.pdf
nsd <- tribble(~ class, ~ total_species, 
               "Echinoidea", 940,
               "Holothuroidea", 1430,
               "Insecta", 1e6,
               "Aves",  9990,
               "Mammalia", 5487,
               "Clitellata", 8e3, # wikipedia
               "Amphibia", 6515,
               "Demospongiae", 8800, # wikipedia
               "Teleostei", 30e3, # wikipedia
               "Bivalvia", 10e3,  # https://ucmp.berkeley.edu/taxa/inverts/mollusca/bivalvia.php
               "Malacostraca", 40e3, # wikipedia
               "Branchiopoda", 800, # animaldiversity.org
               "Gastropoda", prod(c(40e3, 100e3))^(1/2), # https://ucmp.berkeley.edu/taxa/inverts/mollusca/gastropoda.php
               "Chromadorea", 2124, # ITIS
               "Reptilia", 8734,
               "Ascidiacea", 2300, # wikipedia
               "Anthozoa", 6e3, # wikipedia
               "Euchelicerata", 77000, # wikipedia
               "Holostei", 8, # wikipedia
               "Anopla", 600, # wikipedia
               "Arachnida", 102248,
               "Ophiuroidea",  2064, # https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0031940#
               "Polychaeta", 10000, # https://animaldiversity.org/accounts/Polychaeta/
               "Cephalopoda", 800, # https://ucmp.berkeley.edu/taxa/inverts/mollusca/cephalopoda.php
             "Maxillopoda", 14e3 # https://www.sciencedirect.com/topics/agricultural-and-biological-sciences/maxillopoda
)

@@

nsd %>% 
  dt_dps %>% 
  select(phylum, species) %>% group_by(phylum)

# by phyla
dnsp_phyla <- dt_dps %>% 
  filter(is.finite(log10_popsize), is.finite(log10_body_mass_pred)) %>%
  select(phylum, species) %>% group_by(phylum) %>% 
               summarize(nspecies=n())

# by class
dnsp_class <- dt_dps %>% 
  filter(is.finite(log10_popsize), is.finite(log10_body_mass_pred)) %>%
  select(class, species) %>% group_by(class) %>% 
               summarize(nspecies=n())

@@

@@

baron_mine %>% left_join(dnsp_phyla, by='phylum')  %>% select(nspecies.x, nspecies.y)
  mutate(a=signif(total_carbon_mass / frac, 2)) %>% 
  select(-bd_total_biomass) %>%
  select(-total_log10_popsize, -nspecies.y) %>% as.data.frame

@@

How are the proportions?

@@

Ok, now let's look at total biomass. Since I have drastically different phylum
represented here in my taxa sample than on earth, we need to reweight things.

# using the Bar-On weights
# prop_carbon_mass is prop in my dataset
baron_mine <- baron_mine %>% 
              mutate(reweight = bd_prop_mass / prop_carbon_mass) %>%
              mutate(reweighted_total_carbon_mass = reweight * total_carbon_mass)

total_sample_carbon_mass <- baron_mine %>% pull(reweighted_total_carbon_mass) %>% sum()

# total species here
n <- nrow(xd)
total_animal_species <- sum(nspecies)

print(paste0("fraction of animal species in this data set: ", 
             signif(n/total_animal_species, 2)))

# Bar-On et al. (2018) estimate is 2Gt C
# 1e12 kg = 1Gt
total_animal_cmass <- 2

# if the species here were random
signif(total_animal_cmass * n / total_animal_species, 2)
signif(total_sample_carbon_mass, 2)

0.058 / total_animal_cmass

Note: initially it was my plan to see if we could get a sense of what fraction
of the total biomass this dataset contains (which is about 3%). This seems
unreasonably high, even after reweighting as above. The problem is that the
3.4e-5 fraction of animal species is obviously not a random sample based on
size / population, and includes humans, domesticated species, etc. -- this
probably is accentuated by the overabundance of chordates compared arthropods
here.

>>>>

# main sources:
# https://www.environment.gov.au/system/files/pages/2ee3f4a1-f130-465b-9c7a-79373680a067/files/nlsaw-2nd-complete.pdf
# this is for the classes *in my dataset*

geomean <- function(x) (prod(x))^(1/length(x))

nsd <- tribble(~ class, ~ total_species, 
               "Echinoidea", 940,
               "Holothuroidea", 1430,
               "Insecta", 1e6,
               "Aves",  9990,
               "Mammalia", 5487,
               "Clitellata", 8e3, # wikipedia
               "Amphibia", 6515,
               "Demospongiae", 8800, # wikipedia
               "Teleostei", 30e3, # wikipedia
               "Bivalvia", 10e3,  # https://ucmp.berkeley.edu/taxa/inverts/mollusca/bivalvia.php
               "Malacostraca", 40e3, # wikipedia
               "Branchiopoda", 800, # animaldiversity.org
               "Gastropoda", geomean(c(40e3, 100e3)), # https://ucmp.berkeley.edu/taxa/inverts/mollusca/gastropoda.php
               "Chromadorea", 2124, # ITIS
               "Reptilia", 8734,
               "Ascidiacea", 2300, # wikipedia
               "Anthozoa", 6e3, # wikipedia
               "Euchelicerata", 77000, # wikipedia
               "Holostei", 8, # wikipedia
               "Anopla", 600, # wikipedia
               "Arachnida", 102248,
               "Ophiuroidea",  2064, # https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0031940#
               "Polychaeta", 10000, # https://animaldiversity.org/accounts/Polychaeta/
               "Cephalopoda", 800, # https://ucmp.berkeley.edu/taxa/inverts/mollusca/cephalopoda.php
             "Maxillopoda", 14e3 # https://www.sciencedirect.com/topics/agricultural-and-biological-sciences/maxillopoda
)

@@

# what are the total species by phyla? we join in the mapping from our data
total_species_by_phyla <- nsd %>% right_join(phyla_class, by='class') %>% 
  group_by(phylum) %>% summarize(total_species = sum(total_species, na.rm=TRUE))

# these numbers are conditioning on classes we have in this study
# we actually want all classes on *earth*; for this I use the table 
# here from wikipedia
# https://en.wikipedia.org/wiki/Animal#Numbers_and_habitats
total_species_by_phyla2 <- tribble(~ phylum, ~ total_species,
                                   "Annelida", 17e3,
                                   "Arthropoda", 1.257e6,
                                   "Chordata", geomean(c(45e3, 65e3)),
                                   "Cnidaria", 16e3,
                                   "Echinodermata", 7.5e3,
                                   "Mollusca", geomean(c(85e3, 107e3)),
                                   "Nematoda", 25e3,
                                   "Nemertea", 1300,
                                   "Porifera", 10800)

# and by phyla? again joining in by the mapping in our data
# conditioning on classes in my dataset
total_species_by_class <- nsd %>% right_join(phyla_class, by='class') %>% 
  group_by(class) %>% summarize(total_species = sum(total_species, na.rm=TRUE))

# how different are the number of species when we compare the dataset 
# further up, conditioning on classes in my dataset, versus the dataset
# immediately above, from Wikipedia?
total_species_by_phyla %>% left_join(total_species_by_phyla2, by='phylum') %>%
  mutate(ratio = total_species.x / total_species.y)

Nothing terribly off here; nematodes is the only sizeable difference.

```{r}
## now join in the total species in our dataset
# by phyla
dnsp_phyla <- dt_dps %>% 
  left_join(total_species_by_phyla2) %>%
  filter(is.finite(log10_popsize), is.finite(log10_body_mass_pred)) %>%
  select(phylum, species, total_species) %>% group_by(phylum, total_species) %>% 
  summarize(nspecies=n()) %>% 
  mutate(frac = nspecies / total_species)

# by class
dnsp_class <- dt_dps %>% 
  left_join(total_species_by_class) %>%
  filter(is.finite(log10_popsize), is.finite(log10_body_mass_pred)) %>%
  select(phylum, class, species, total_species) %>% group_by(phylum, class, total_species) %>% 
  summarize(nspecies=n()) %>%
  mutate(frac = nspecies / total_species)

@@

# this would make a cool supp plot, saving data
write_tsv(dnsp_class, file='../data/species_by_class.tsv')

@@

# join everything
biomass <- baron_mine %>% left_join(dnsp_phyla, by='phylum')  %>% 
  # 
  mutate(factor=signif(total_carbon_mass / (bd_mass * frac), 3)) %>% 
  select(-bd_total_biomass) %>%
  select(-total_log10_popsize)

stopifnot(all(biomass$nspecies.x == biomass$nspecies.y))
biomass <- biomass %>% rename(nspecies = nspecies.x) %>%
  select(-nspecies.y, -total)

write_tsv(biomass, file='../data/biomass.tsv')

bm_tbl <- biomass %>% mutate(total_species = as.integer(total_species)) %>%
           select(phylum, total_species, bd_mass, bd_prop_mass, 
                  total_carbon_mass, prop_carbon_mass, nspecies, frac, factor)

print(xtable(bm_tbl, digits=c(1, 2, -2, 2, 4, -2, 4, 2, -2, 2)), 
      include.rownames=FALSE,
      tabular.environment='tabular',
      math.style.exponents=TRUE, type='latex')

@@

Things look relatively good here -- nothing is off by several orders of
magnitude unless there's only 1-3 species sampled.

How are the proportions of biomass by taxa?

@@
